---
title: "Generate heatmaps of TF activity across tissues and barplots of TF tissue specificity"
author: "Sharan Surya"
date: "`r Sys.Date()`"
format:
  html:
    toc: true
    toc-depth: 3
    code-fold: show
    code-tools: true
    theme: cosmo
    fig-width: 12
    fig-height: 9
    embed-resources: true
---

# Overview

```{r setup, include=FALSE}

rm(list = ls())

knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  fig.align = "center",
  dpi = 300
)

options(scipen = 10000)

```

# Load Required Packages

```{r}
#| label: setup-libraries
#| message: false
#| warning: false

library(ggplot2)
library(dplyr)
library(tidyverse)
library(patchwork)
library(seriation)
```

# Load Data

```{r load-log2fc-data}
# Load the combined log2FC file
log2fc_data <- readRDS("input/all_genes_per_tissue_log2FC_20231116.rds")

# Filter to protein-coding genes only
log2fc_data <- log2fc_data[log2fc_data$gene_biotype == "protein_coding", ]
all_tf_activity <- readRDS("celest/output/all_tf_activity.rds")

```


# Heatmap of predicted TFs for each tissue

```{r tf-heatmap-prep}
# Fixed column order for all heatmaps
tissue_order <- c("Germline", "SGP", "hmc", "GLR", "Pblast", "NonStriatedMuscle", "BWM", "Excretory", "Pharynx", "NonHypodermEpithelium", "Coelomocyte", "Glia", "NonSeamHypoderm", "Seam", "PharyngealNeuron", "Neuron", "CiliatedNeuron", "Intestine")

# Combine all tissue TF activity data into one long data frame
tf_activity_long <- bind_rows(all_tf_activity) %>%
  select(wb_id, public_name, score, p_value, tissue)

# Get union of all TFs across tissues
all_tfs <- unique(tf_activity_long$public_name)
```

## Heatmap 1: TF Activity Score per Tissue

```{r tf-heatmap-score, fig.width=14, fig.height=20}
# Build TF x tissue matrix for OLO computation
score_mat <- tf_activity_long %>%
  select(public_name, tissue, score) %>%
  pivot_wider(names_from = tissue, values_from = score) %>%
  column_to_rownames("public_name") %>%
  as.matrix()

score_mat <- score_mat[, tissue_order[tissue_order %in% colnames(score_mat)]]
score_mat[is.na(score_mat)] <- 0

# OLO row ordering
dist_rows <- dist(score_mat, method = "euclidean")
olo_order <- seriation::seriate(dist_rows, method = "OLO")
tf_levels <- rownames(score_mat)[seriation::get_order(olo_order)]

# To ensure top is red (positive activity) - check if top TFs have higher mean score than bottom TFs; if not, reverse the order
tf_means <- rowMeans(score_mat[tf_levels, , drop = FALSE], na.rm = TRUE)
if (mean(head(tf_means, 10)) > mean(tail(tf_means, 10))) {
  tf_levels <- rev(tf_levels)
} # why ">"? because ggplot plots from bottom to top.

# Back to long format with factor levels set by OLO
score_long <- tf_activity_long %>%
  select(public_name, tissue, score) %>%
  mutate(
    public_name = factor(public_name, levels = tf_levels),
    tissue      = factor(tissue, levels = tissue_order)
  )

heatmap_score <- ggplot(score_long, aes(x = tissue, y = public_name, fill = score)) +
  geom_tile(color = NA) +
  scale_fill_gradient2(
    low      = "#2C7BB6",
    mid      = "white",
    high     = "#D7191C",
    midpoint = 0,
    na.value = "grey90",
    name     = "Activity\nScore"
  ) +
  labs(
    title = "TF Activity Score Across Tissues",
    x     = "Tissue",
    y     = NULL
  ) +
  theme_minimal(base_size = 11) +
  theme(
    plot.title      = element_text(size = 14, face = "bold", hjust = 0.5),
    axis.text.x     = element_text(angle = 45, hjust = 1, size = 10),
    axis.text.y     = element_blank(),
    axis.ticks.y    = element_blank(),
    panel.grid      = element_blank(),
    legend.position = "right"
  )

ggsave(
  filename = "celest/plots/tf_heatmap_score.pdf", 
  plot = heatmap_score,
  width = 12, 
  height = 12, 
  units = "in",
  device = cairo_pdf
)


```

## Heatmap 2: TF p-value per Tissue

```{r tf-heatmap-pvalue, fig.width=14, fig.height=20}
# Build -log10(p_value) matrix for OLO computation
pval_mat <- tf_activity_long %>%
  select(public_name, tissue, p_value) %>%
  mutate(neg_log10_pval = -log10(p_value + 1e-300)) %>%
  select(public_name, tissue, neg_log10_pval) %>%
  pivot_wider(names_from = tissue, values_from = neg_log10_pval) %>%
  column_to_rownames("public_name") %>%
  as.matrix()

pval_mat <- pval_mat[, tissue_order[tissue_order %in% colnames(pval_mat)]]

# OLO row ordering
dist_rows_pval <- dist(pval_mat, method = "euclidean")

dist_rows_pval[is.na(dist_rows_pval)] <- 0

olo_order_pval <- seriation::seriate(dist_rows_pval, method = "OLO")
tf_levels_pval <- rownames(pval_mat)[seriation::get_order(olo_order_pval)]

# To ensure top is red (positive activity) - check if top TFs have higher mean score than bottom TFs; if not, reverse the order
tf_means_pval <- rowMeans(pval_mat[tf_levels_pval, , drop = FALSE], na.rm = TRUE)
if (mean(head(tf_means_pval, 10)) > mean(tail(tf_means_pval, 10))) {
  tf_levels_pval <- rev(tf_levels_pval)
}

# Back to long format
pval_long <- tf_activity_long %>%
  select(public_name, tissue, p_value) %>%
  mutate(
    neg_log10_pval = -log10(p_value + 1e-300),
    public_name    = factor(public_name, levels = tf_levels_pval),
    tissue         = factor(tissue, levels = tissue_order)
  )

heatmap_pval <- ggplot(pval_long, aes(x = tissue, y = public_name, fill = neg_log10_pval)) +
  geom_tile(color = NA) +
  scale_fill_gradient2(
    low      = "#2C7BB6",
    mid      = "white",
    high     = "#D7191C",
    midpoint = 0,
    na.value = "grey90",
    name     = "-log10\np-value"
  ) +
  labs(
    title = "TF Significance Across Tissues  (-log10 p-value)",
    x     = "Tissue",
    y     = NULL
  ) +
  theme_minimal(base_size = 11) +
  theme(
    plot.title      = element_text(size = 14, face = "bold", hjust = 0.5),
    axis.text.x     = element_text(angle = 45, hjust = 1, size = 10),
    axis.text.y     = element_blank(),
    axis.ticks.y    = element_blank(),
    panel.grid      = element_blank(),
    legend.position = "right"
  )


ggsave(
  filename = "celest/plots/tf_heatmap_pvalue.pdf", 
  plot = heatmap_pval,
  width = 12, 
  height = 12, 
  units = "in",
  device = cairo_pdf
)

```

## Heatmap 3: Score (significant TFs only, p-value < 0.05)

```{r tf-heatmap-score-sig, fig.width=14, fig.height=20}
# Mask score with NA where p_value >= 0.05
sig_long <- tf_activity_long %>%
  mutate(score_sig = if_else(p_value < 0.05, score, NA_real_))


# Keep only TFs significant in at least one tissue
tfs_with_sig <- sig_long %>%
  group_by(public_name) %>%
  summarise(n_sig = sum(!is.na(score_sig)), .groups = "drop") %>%
  filter(n_sig >= 1) %>%
  pull(public_name)

tfs_with_sig_wbid <- sig_long %>%
  group_by(wb_id) %>%
  summarise(n_sig = sum(!is.na(score_sig)), .groups = "drop") %>%
  filter(n_sig >= 1) %>%
  pull(wb_id)

write_delim(
  data.frame(tfs_with_sig_wbid),
  "celest/output/tf_with_sig_wbid.txt",
  delim = ",",
  col_names = FALSE
)
# Build matrix for OLO (NAs replaced with 0 only for distance calculation)
sig_mat <- sig_long %>%
  filter(public_name %in% tfs_with_sig) %>%
  select(public_name, tissue, score_sig) %>%
  pivot_wider(names_from = tissue, values_from = score_sig) %>%
  column_to_rownames("public_name") %>%
  as.matrix()

sig_mat <- sig_mat[, tissue_order[tissue_order %in% colnames(sig_mat)]]

sig_mat_nona <- sig_mat
sig_mat_nona[is.na(sig_mat_nona)] <- 0
dist_rows_sig <- dist(sig_mat_nona, method = "euclidean")
olo_order_sig <- seriation::seriate(dist_rows_sig, method = "OLO")
tf_levels_sig <- rownames(sig_mat)[seriation::get_order(olo_order_sig)]

# To ensure top is red (positive activity) - check if top TFs have higher mean score than bottom TFs; if not, reverse the order
tf_means_sig <- rowMeans(sig_mat[tf_levels_sig, , drop = FALSE], na.rm = TRUE)
if (mean(head(tf_means_sig, 10)) > mean(tail(tf_means_sig, 10))) {
  tf_levels_sig <- rev(tf_levels_sig)
}

# Back to long format, with OLO factor levels
sig_long_filtered <- sig_long %>%
  filter(public_name %in% tfs_with_sig) %>%
  complete(public_name, tissue = tissue_order) %>%   # creates missing rows as NA
  mutate(
    public_name = factor(public_name, levels = tf_levels_sig),
    tissue      = factor(tissue, levels = tissue_order)
  )

heatmap_score_sig <- ggplot(sig_long_filtered,
                             aes(x = tissue, y = public_name, fill = score_sig)) +
  geom_tile(color = NA) +
  scale_fill_gradient2(
    low      = "#2C7BB6",
    mid      = "white",
    high     = "#D7191C",
    midpoint = 0,
    na.value = "grey90",
    name     = "Activity\nScore"
  ) +
  labs(
    title    = "TF Activity Score Across Tissues (p < 0.05 only)",
    subtitle = paste0(length(tfs_with_sig),
                      " TFs with \u22651 significant tissue shown; grey = not significant"),
    x        = "Tissue",
    y        = NULL
  ) +
  theme_minimal(base_size = 11) +
  theme(
    aspect.ratio = 0.7,
    plot.title      = element_text(size = 14, face = "bold", hjust = 0.5),
    plot.subtitle   = element_text(size = 10, hjust = 0.5, color = "grey40"),
    axis.text.x     = element_text(angle = 45, hjust = 1, size = 10),
    axis.text.y     = element_blank(),
    axis.ticks.y    = element_blank(),
    panel.grid      = element_blank(),
    legend.position = "right"
  )

ggsave(
  filename = "celest/plots/tf_heatmap_score_significant.pdf", 
  plot = heatmap_score_sig,
  width = 12, 
  height = 12, 
  units = "in",
  device = cairo_pdf
)

```


## Heatmap 4: Score (significant TFs only, p-value < 0.05) top50

```{r tf-heatmap-score-sig-top50, fig.width=14, fig.height=20}
# Mask score with NA where p_value >= 0.05
sig_long_t50 <- tf_activity_long %>%
  mutate(score_sig = if_else(p_value < 0.05, score, NA_real_))

write_delim(
  data.frame(unique(sig_long_t50$wb_id)),
  "celest/output/all_tfs_wbid.txt",
  delim = ",",
  col_names = FALSE
)

# Rank TFs by number of significant tissues, then mean score; keep top 50
tf_order_top50 <- sig_long_t50 %>%
  group_by(public_name) %>%
  summarise(
    n_sig      = sum(!is.na(score_sig)),
    mean_score = mean(score_sig, na.rm = TRUE),
    .groups    = "drop"
  ) %>%
  filter(n_sig >= 1) %>%
  arrange(desc(n_sig), desc(mean_score)) %>%
  slice_head(n = 50) %>%
  pull(public_name)

tf_order_top50_wbid <- sig_long_t50 %>%
  group_by(wb_id) %>%
  summarise(
    n_sig      = sum(!is.na(score_sig)),
    mean_score = mean(score_sig, na.rm = TRUE),
    .groups    = "drop"
  ) %>%
  filter(n_sig >= 1) %>%
  arrange(desc(n_sig), desc(mean_score)) %>%
  slice_head(n = 50) %>%
  pull(wb_id)

write_delim(
  data.frame(tf_order_top50_wbid),
  "celest/output/tf_order_top50_wbid.txt",
  delim = ",",
  col_names = FALSE
)

# Build top-50 matrix for OLO
sig_mat_t50 <- sig_long_t50 %>%
  filter(public_name %in% tf_order_top50) %>%
  select(public_name, tissue, score_sig) %>%
  pivot_wider(names_from = tissue, values_from = score_sig) %>%
  column_to_rownames("public_name") %>%
  as.matrix()

sig_mat_t50 <- sig_mat_t50[, tissue_order[tissue_order %in% colnames(sig_mat_t50)]]

sig_mat_t50_nona <- sig_mat_t50
sig_mat_t50_nona[is.na(sig_mat_t50_nona)] <- 0
dist_rows_t50 <- dist(sig_mat_t50_nona, method = "euclidean")
olo_order_t50 <- seriation::seriate(dist_rows_t50, method = "OLO")
tf_levels_t50 <- rownames(sig_mat_t50)[seriation::get_order(olo_order_t50)]

tf_means_t50 <- rowMeans(sig_mat_t50[tf_levels_t50, , drop = FALSE], na.rm = TRUE)
if (mean(head(tf_means_t50, 10)) > mean(tail(tf_means_t50, 10))) {
  tf_levels_t50 <- rev(tf_levels_t50)
}

# Back to long format with OLO factor levels
sig_long_t50_filtered <- sig_long_t50 %>%
  filter(public_name %in% tf_order_top50) %>%
  complete(public_name, tissue = tissue_order) %>%   # creates missing rows as NA
  mutate(
    public_name = factor(public_name, levels = tf_levels_t50),
    tissue      = factor(tissue, levels = tissue_order)
  )

heatmap_score_t50 <- ggplot(sig_long_t50_filtered,
                             aes(x = tissue, y = public_name, fill = score_sig)) +
  geom_tile(color = NA) +
  scale_fill_gradient2(
    low      = "#2C7BB6",
    mid      = "white",
    high     = "#D7191C",
    midpoint = 0,
    na.value = "grey90",
    name     = "Activity\nScore"
  ) +
  labs(
    title    = "TF Activity Score Across Tissues (p < 0.05 only, top 50 TFs)",
    subtitle = "Top 50 TFs by tissue breadth and mean score; grey = not significant",
    x        = "Tissue",
    y        = NULL
  ) +
  theme_minimal(base_size = 11) +
  theme(
    plot.title      = element_text(size = 14, face = "bold", hjust = 0.5),
    plot.subtitle   = element_text(size = 10, hjust = 0.5, color = "grey40"),
    axis.text.x     = element_text(angle = 45, hjust = 1, size = 12),
    axis.text.y     = element_text(size = 12),
    axis.ticks.y    = element_blank(),
    panel.grid      = element_blank(),
    legend.position = "right"
  )

ggsave(
  filename = "celest/plots/tf_heatmap_score_significant_t50.pdf", 
  plot = heatmap_score_t50,
  width = 12, 
  height = 12, 
  units = "in",
  device = cairo_pdf
)

```

## Heatmap 5: Score (significant TFs only, p-value < 0.05) top50 based on highest score, not no. of tissues

```{r tf-heatmap-score-sig-top50-score, fig.width=14, fig.height=20}

# Rank TFs by max absolute score; keep top 50

sig_long_t50_score <- tf_activity_long %>%
  mutate(score_sig = if_else(p_value < 0.05, score, NA_real_))

tf_rank_top50_score <- sig_long_t50_score %>%
  filter(!is.na(score_sig)) %>%          # drop NA rows upfront
  group_by(public_name) %>%
  summarise(
    score_sig = score_sig[which.max(abs(score_sig))],
    .groups = "drop"
  ) %>%
  filter(is.finite(score_sig)) %>%
  arrange(desc(abs(score_sig))) %>%
  slice_head(n = 50)

tf_order_top50_score <- tf_rank_top50_score$public_name

write.table(tf_rank_top50_score,
            file = "celest/output/tf_rank_top50_score.tsv",
            sep = "\t",
            row.names = FALSE,
            quote = FALSE)

tf_rank_top50_score_wbid <- sig_long_t50_score %>%
  filter(!is.na(score_sig)) %>%          # drop NA rows upfront
  group_by(wb_id) %>%
  summarise(
    max_score_sig = max(abs(score_sig), na.rm = TRUE),
    .groups       = "drop"
  ) %>%
  filter(is.finite(max_score_sig)) %>%
  arrange(desc(max_score_sig)) %>%
  slice_head(n = 50) %>%
  pull(wb_id)

write_delim(
  data.frame(tf_rank_top50_score_wbid),
  "celest/output/tf_rank_top50_score_wbid.txt",
  delim = ",",
  col_names = FALSE
)

# Build top-50 matrix for OLO

sig_mat_t50_score <- sig_long_t50_score %>%
  filter(public_name %in% tf_order_top50_score) %>%
  select(public_name, tissue, score_sig) %>%
  pivot_wider(names_from = tissue, values_from = score_sig) %>%
  column_to_rownames("public_name") %>%
  as.matrix()

# Ensure tissue order
sig_mat_t50_score <- sig_mat_t50_score[
  , tissue_order[tissue_order %in% colnames(sig_mat_t50_score)]
]

# Replace NA with 0 for distance computation
sig_mat_t50_score_nona <- sig_mat_t50_score
sig_mat_t50_score_nona[is.na(sig_mat_t50_score_nona)] <- 0


dist_rows_t50_score <- dist(sig_mat_t50_score_nona, method = "euclidean")
olo_order_t50_score <- seriation::seriate(dist_rows_t50_score, method = "OLO")

tf_levels_t50_score <- rownames(sig_mat_t50_score)[
  seriation::get_order(olo_order_t50_score)
]
  

tf_means_t50_score <- rowMeans(sig_mat_t50_score[tf_levels_t50_score, , drop = FALSE], na.rm = TRUE)
if (mean(head(tf_means_t50_score, 10)) < mean(tail(tf_means_t50_score, 10))) {
  tf_levels_t50_score <- rev(tf_levels_t50_score)
}

# Back to long format with OLO factor levels
sig_long_t50_filtered_score <- sig_long_t50_score %>%
  filter(public_name %in% tf_order_top50_score) %>%
  complete(public_name, tissue = tissue_order) %>%   # creates missing rows as NA
  mutate(
    public_name = factor(public_name, levels = rev(tf_levels_t50_score)),
    tissue      = factor(tissue, levels = tissue_order)
  )

heatmap_score_t50_score <- ggplot(
  sig_long_t50_filtered_score,
  aes(x = tissue, y = public_name, fill = score_sig)
) +
  geom_tile(color = NA) +
  scale_fill_gradient2(
    low      = "#2166AC",
    mid      = "white",
    high     = "#B2182B",
    midpoint = 0,
    na.value = "grey90",
    name     = "Activity\nScore"
  ) +
  labs(
    x        = NULL,
    y        = NULL
  ) +
  theme_minimal(base_size = 8) +
  theme(
    axis.title      = element_text(face = "bold"),
    axis.text.x     = element_text(angle = 45, hjust = 1, size = 8),
    axis.text.y     = element_text(size = 6),
    axis.ticks.y    = element_blank(),
    panel.grid      = element_blank(),
    legend.position = "right"
  )

ggsave(
  filename = "celest/plots/tf_heatmap_score_significant_t50_score.pdf", 
  plot = heatmap_score_t50_score,
  width = 6.5, 
  height = 5, 
  units = "in",
  device = cairo_pdf
)
```

# Number of TFs Enriched per Number of Tissues

How many TFs are significantly active (p < 0.05) in exactly 1, 2, 3 … 18 tissues?

```{r tf-tissue-breadth-barplot, fig.width=10, fig.height=6}
# Count number of significant tissues per TF
tf_breadth <- tf_activity_long %>%
  filter(p_value < 0.05) %>%
  group_by(public_name) %>%
  summarise(n_tissues = n_distinct(tissue), .groups = "drop")

# Count how many TFs fall into each tissue-count bin; ensure all 1–18 are present
tf_breadth_counts <- tf_breadth %>%
  dplyr::count(n_tissues, name = "n_tfs") %>%
  complete(n_tissues = 1:18, fill = list(n_tfs = 0))

barplot_breadth <- ggplot(tf_breadth_counts, aes(x = n_tissues, y = n_tfs)) +
  geom_col(fill = "#3498DB", colour = "white", width = 0.75, alpha = 0.8) +
  geom_text(
    aes(label = ifelse(n_tfs > 0, n_tfs, "")),
    vjust = -0.4,
    size  = 3.5,
    fontface = "bold",
    colour = "#2C3E50"
  ) +
  scale_x_continuous(
    breaks = 1:18,
    limits = c(0.25, 18.75),
    expand = c(0, 0)
  ) +
  scale_y_continuous(
    expand = expansion(mult = c(0, 0.12))
  ) +
  labs(
    title    = "Number of TFs Enriched Across Tissues",
    subtitle = "Significance threshold: p < 0.05  |  Total tissues = 18",
    x        = "Number of tissues TF is significantly active in",
    y        = "Number of TFs"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    plot.title    = element_text(size = 14, face = "bold",  hjust = 0.5),
    plot.subtitle = element_text(size = 10, colour = "grey40", hjust = 0.5),
    axis.title    = element_text(size = 11, face = "bold"),
    axis.text     = element_text(size = 10),
    panel.grid.major.x = element_blank(),
    panel.grid.minor   = element_blank(),
    panel.border  = element_rect(colour = "black", fill = NA, linewidth = 0.7)
  )

ggsave(
  filename = "celest/plots/tf_tissue_breadth_barplot.pdf", 
  plot = barplot_breadth,
  width = 6.5, 
  height = 5, 
  units = "in",
  device = cairo_pdf
)

```


# Number of TFs Enriched per Tissue (with Direction of Activity)

How many TFs are significantly active (p < 0.05) in each tissue, and in which direction?

```{r tf-tissue-direction-barplot, fig.width=12, fig.height=7}
# Collect significant TFs with direction of activity
sig_tf_direction <- lapply(all_tf_activity, function(x) {
  x_sig <- x[x$p_value < 0.05, c("public_name", "score", "tissue")]
  x_sig
})

# Combine all tissues into one data frame
sig_tf_direction_df <- do.call(rbind, sig_tf_direction)

# Add direction column
sig_tf_direction_df$direction <- ifelse(
  sig_tf_direction_df$score > 0, "Up",
  ifelse(sig_tf_direction_df$score < 0, "Down", NA)
)

# Remove any NA directions (just in case score == 0)
sig_tf_direction_df <- sig_tf_direction_df[!is.na(sig_tf_direction_df$direction), ]

# Count number of TFs per tissue per direction
tf_tissue_dir_counts <- sig_tf_direction_df %>%
  group_by(tissue, direction) %>%
  summarise(n_tfs = n_distinct(public_name), .groups = "drop") %>%
  mutate(direction = factor(direction, levels = c("Up", "Down")))

# Compute total (Up + Down) per tissue for sorting
tissue_totals <- tf_tissue_dir_counts %>%
  group_by(tissue) %>%
  summarise(total = sum(n_tfs), .groups = "drop") %>%
  mutate(tissue = factor(tissue, levels = tissue_order))# if wanted to arrange tissues in terms of No. of TFs,change it to -> arrange(desc(total))

tf_tissue_dir_counts <- tf_tissue_dir_counts %>%
  mutate(tissue = factor(tissue, levels = tissue_order))


barplot_direction <- ggplot(tf_tissue_dir_counts,
                            aes(x = tissue, y = n_tfs, fill = direction)) +
  geom_col(width = 0.75, alpha = 1, position = "stack") +
  geom_text(
    data = tissue_totals,
    aes(x = tissue, y = total, label = total),
    vjust       = -0.4,
    size        = 2,
    fontface    = "bold",
    colour      = "#000000",
    inherit.aes = FALSE
  ) +
  scale_fill_manual(
    values = c("Up" = "#B2182B", "Down" = "#2166AC"),
    name   = "Activity"
  ) +
  scale_y_continuous(
    expand = expansion(mult = c(0, 0.12))
  ) +
  labs(
    x        = "Tissue",
    y        = "Number of TFs"
  ) +
  theme_classic(base_size = 8) +
  theme(
    axis.title    = element_text(size = 8, face = "bold"),
    axis.text.x   = element_text(angle = 45, hjust = 1),
    legend.position = "top",
    legend.key.size = unit(0.2, "cm")
  )

ggsave(
  filename = "celest/plots/tf_tissue_direction_barplot.pdf", 
  plot = barplot_direction,
  width = 3.2, 
  height = 3, 
  units = "in",
  device = cairo_pdf
)
```

# Number of TFs Enriched per Number of Tissues (with Direction of Activity)

How many TFs are significantly active (p < 0.05) in exactly 1, 2, 3 … 18 tissues, broken down by direction of activity?

```{r tf-breadth-direction-barplot, fig.width=10, fig.height=6}
# Count number of significant tissues per TF, with direction
tf_breadth_dir <- sig_tf_direction_df %>%
  group_by(public_name, direction) %>%
  summarise(n_tissues = n_distinct(tissue), .groups = "drop")

# Count how many TFs fall into each tissue-count bin per direction
tf_breadth_dir_counts <- tf_breadth_dir %>%
  group_by(n_tissues, direction) %>%
  summarise(n_tfs = n(), .groups = "drop") %>%
  complete(n_tissues = 1:18, direction = c("Up", "Down"), fill = list(n_tfs = 0)) %>%
  mutate(direction = factor(direction, levels = c("Up", "Down")))

barplot_breadth_dir <- ggplot(tf_breadth_dir_counts,
                              aes(x = n_tissues, y = n_tfs, fill = direction)) +
  geom_col(width = 0.75, alpha = 0.85, position = "stack") +
  geom_text(
    data = tf_breadth_dir_counts %>%
      group_by(n_tissues) %>%
      summarise(total = sum(n_tfs), .groups = "drop") %>%
      filter(total > 0),
    aes(x = n_tissues, y = total, label = total),
    vjust    = -0.4,
    size     = 2,
    fontface = "bold",
    colour   = "#000000",
    inherit.aes = FALSE
  ) +
  scale_x_continuous(
    breaks = 1:18,
    limits = c(0.25, 18.75),
    expand = c(0, 0)
  ) +
  scale_y_continuous(
    expand = expansion(mult = c(0, 0.12))
  ) +
  scale_fill_manual(
    values = c("Up" = "#B2182B", "Down" = "#2166AC"),
    name   = "Activity"
  ) +
  labs(
    x        = "Number of tissues TF is significantly active in",
    y        = "Number of TFs"
  ) +
  theme_classic(base_size = 8) +
  theme(
    axis.title    = element_text(size = 8, face = "bold"),
    axis.text     = element_text(size = 8),
    legend.position = "top",
    legend.key.size = unit(0.2, "cm")
  )


ggsave(
  filename = "celest/plots/tf_tissue_breadth_direction_barplot.pdf", 
  plot = barplot_breadth_dir,
  width = 3.2, 
  height = 3, 
  units = "in",
  device = cairo_pdf
)
```


# TF Activity vs Log2FC across tissues 

```{r tf-activity-log2fc, fig.width=10, fig.height=6}

# merge significant TF activity data with significant log2FC data for the corresponding TFs and tissues

tf_activity_log2fc <- sig_long_filtered %>%
  filter(p_value < 0.05) %>% 
  left_join(log2fc_data, by = c("wb_id" = "gene_id", "tissue" = "tissue")) %>%
  filter(padj < 0.05)

saveRDS(tf_activity_log2fc, "celest/output/tf_activity_log2fc.rds")
```



# Session Info

```{r}
sessionInfo()
```

