---
title: "Transcription Factor Activity Estimation with CelEsT"
author: "Sharan Surya"
format: 
  html:
    embed-resources: true
    toc: true
    toc-depth: 3
    number-sections: true
    code-fold: show
    code-line-numbers: true
    highlight-style: github  
    theme: cosmo
---


# Overview

Transcription factor (TF) activity estimation using the orthCelEsT gene regulatory network and DESeq2 differential expression statistics. The analysis uses the multivariate linear model (MLM) method from the decoupleR package to infer TF activities from gene expression changes.



```{r}
#| label: document-chunk-options
#| include: false

rm(list = ls())

knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  fig.align = "center",
  dpi = 300
)

options(scipen = 10000)

```

# Load Required Packages

```{r load-packages, eval=FALSE}
# Package installation code (already installed)
# packages <- c("stringr", "openxlsx", "edgeR", "splines", 
#               "dplyr", "ggplot2", "ggrepel")
# 
# biocmanager_packages <- c("DESeq2", "decoupleR")
# note - wormRef has RAPToR as dependency, and RAPToR has pryr as dependency. However pryr is not available for R 4.6.1, so install RAPToR locally by editing out and replacing the pryr functions manually. 
```

```{r}
#| label: setup-libraries
#| message: false
#| warning: false

library(openxlsx)
library(DESeq2)
library(decoupleR)
library(wormRef)
library(dplyr)
library(readxl)
library(readr)
library(wbData)
library(tidyr)
library(writexl)
library(stringr)

```

# Define Helper Functions

## Remove Duplicated Genes

This function ensures no duplicated gene IDs are present in the dataset by removing all instances of duplicated IDs.

```{r function-remove-duplicates}
remove.duplicated.genes <- function(geneIDs) {
  if (any(duplicated(geneIDs))) {
    n_duplicates <- sum(duplicated(geneIDs) | duplicated(geneIDs, fromLast = TRUE))
    
    unique_gene_ids <- geneIDs[!(duplicated(geneIDs) | 
                                  duplicated(geneIDs, fromLast = TRUE))]
    
    message(paste0(n_duplicates, 
                   " genes removed from analysis due to duplicated gene IDs"))
    
    return(unique_gene_ids)
  } else {
    return(geneIDs)
  }
}
```

# Load Input Data

## Load TF Annotations and CelEsT Network

The CelEsT network contains transcription factor-target gene interactions with associated weights representing the mode of regulation. Also load the original .rds file to get the tissue names.

```{r load-data}
# Load TF annotations from manuscript supplementary files
tf_annotations <- read_excel("celest/input/Table S1 - TF annotation.xlsx", sheet = 2)

# Load CelEsT gene regulatory network
# For alternative networks (orthCelEsT or maxCelEsT), modify the file path
#celest_network <- read_excel("celest/input/Table S3 - CelEsT network annotated.xlsx", sheet = 2)
celest_network <- read.delim("celest/input/orthCelEsT.txt", header = TRUE, stringsAsFactors = FALSE)

all_genes_per_tissue_log2FC_20231116 <- readRDS("input/all_genes_per_tissue_log2FC_20231116.rds")
colnames(all_genes_per_tissue_log2FC_20231116)
```

## Load Differential Expression Statistics

We'll process all tissue-specific DE statistics files in a loop.


```{r load-de-stats}

tissue_names <- unique(all_genes_per_tissue_log2FC_20231116$tissue)

# Alternative: Auto-detect tissue names from files in Input directory
# tissue_files <- list.files("Input", pattern = "^waldstat_per_tissue_.*\\.txt$", full.names = FALSE)
# tissue_names <- gsub("waldstat_per_tissue_|\\.txt$", "", tissue_files)

cat("Found", length(tissue_names), "tissues to analyze:\n")
cat(paste(tissue_names, collapse = ", "), "\n\n")
```

# Load detection_matrix to filter TFs prior to analysis 
```{r load-detection}
detection_matrix <- read.delim("output/detection_matrix_tissue.txt", header = TRUE, stringsAsFactors = FALSE)

# add column sequence and match gene_id to sequence
gids <- wb_load_gene_ids("WS296")
detection_matrix <- detection_matrix |> 
  left_join(gids |> 
              select(sequence, gene_id),
            by = c("X" = "gene_id"))

```
# Process All Tissues

This section will loop through all tissue-specific DE statistics files and perform TF activity estimation for each tissue.

Note: This can be done without looping! DecoupleR can take a matrix of DE statistics with multiple columns (one per tissue) and will return a matrix of TF activities with the same structure. Did not know this before!

```{r process-all-tissues}

# Initialize list to store results
all_tf_activity <- list()
all_backgrounds <- list()


# Loop through each tissue
for (tissue in tissue_names) {
  
  cat("\n", rep("=", 60), "\n", sep = "")
  cat("Processing:", tissue, "\n")
  cat(rep("=", 60), "\n\n", sep = "")
  
  # Construct file path
  input_file <- paste0("celest/input/waldstat/waldstat_per_tissue_", tissue, ".txt")
  
  # Check if file exists
  if (!file.exists(input_file)) {
    cat("WARNING: File not found:", input_file, "- Skipping\n")
    next
  }
  
  # Filter celest_network to TFs detected in this tissue
  if (!tissue %in% colnames(detection_matrix)) {
    cat("WARNING: Tissue '", tissue, "' not found in detection_matrix - Skipping\n")
    next
  }
  
  detected_tfs <- detection_matrix$sequence[detection_matrix[[tissue]] == 1]
  tissue_network <- celest_network[celest_network$source %in% detected_tfs, ]
  all_backgrounds[[tissue]] <- unique(tissue_network$source)

  cat("  TFs in network:  ", length(unique(celest_network$source)), "\n")
  cat("  TFs detected:    ", length(detected_tfs), "\n")
  cat("  TFs after filter:", length(unique(tissue_network$source)), "\n")
  # ──────────────────────────────────────────────────────────────────────────
  
  # Load differential expression statistics
  de_stats_raw <- read.delim(input_file)
  
  cat("  Loaded", nrow(de_stats_raw), "genes from", input_file, "\n")
  
  # Filter and format gene IDs
  de_stats_filtered <- de_stats_raw[
    str_remove(de_stats_raw[, 1], "\\.[a-z]{1,2}") %in% unlist(Cel_genes), 
  ]
  
  # Remove duplicated genes
  unique_genes <- remove.duplicated.genes(de_stats_filtered[, 1])
  de_stats_filtered <- de_stats_filtered[match(unique_genes, de_stats_filtered[, 1]), ]
  
  # Extract DE statistics column and set gene IDs as row names
  de_stats <- de_stats_filtered[, 2, drop = FALSE]
  rownames(de_stats) <- unique_genes
  
  # Replace N A values with 0
  de_stats[is.na(de_stats)] <- 0
  
  # Convert to sequence names
  de_stats$sequence_name <- Cel_genes[
    match(rownames(de_stats), Cel_genes[, "wb_id"]), 
    "sequence_name"
  ]
  
  # Remove duplicates in sequence names
  unique_seq_names <- remove.duplicated.genes(de_stats[, 2])
  de_stats <- de_stats[match(unique_seq_names, de_stats[, 2]), ]
  
  # Set sequence names as row names and keep only statistics column
  rownames(de_stats) <- unique_seq_names
  de_stats <- de_stats[, 1, drop = FALSE]
  
  cat("  Final matrix:", nrow(de_stats), "genes\n")
  
  # Estimate TF activity using the tissue-filtered network
  cat("  Estimating TF activity...\n")
  
  tf_activity <- decoupleR::decouple(
    mat = de_stats[, 1, drop = FALSE], 
    network = tissue_network,          # <-- filtered network used here
    .source = "source",
    .target = "target",
    statistics = "mlm",
    args = list(mlm = list(.mor = "weight")),
    consensus_score = FALSE
  )
  
  cat("  Analyzed", nrow(tf_activity), "TFs\n")
  
  # Correct p-values for multiple testing
  tf_activity$p_value <- p.adjust(tf_activity$p_value, method = "BH")
  
  # Annotate results
  tf_activity <- cbind(
    tf_activity,
    Cel_genes[match(tf_activity$source, Cel_genes$sequence_name), c("wb_id", "public_name")]
  )
  
  tf_activity$public_name <- toupper(tf_activity$public_name)
  
  tf_activity <- tf_activity %>%
    left_join(
      tf_annotations %>%
        select(wormbase_gseq, Wormbase_mode_of_regulation, Uniprot_mode_of_regulation),
      by = c("source" = "wormbase_gseq")
    ) %>%
    mutate(
      Wormbase_mode_of_regulation = case_when(
        Wormbase_mode_of_regulation == -1 ~ "REPRESSOR",
        Wormbase_mode_of_regulation ==  0 ~ "BIFUNCTIONAL",
        Wormbase_mode_of_regulation ==  1 ~ "ACTIVATOR",
        TRUE ~ NA_character_
      ),
      Uniprot_mode_of_regulation = case_when(
        Uniprot_mode_of_regulation == -1 ~ "REPRESSOR",
        Uniprot_mode_of_regulation ==  0 ~ "BIFUNCTIONAL",
        Uniprot_mode_of_regulation ==  1 ~ "ACTIVATOR",
        TRUE ~ NA_character_
      )
    )
  
  # Sort by significance
  tf_activity <- tf_activity[order(tf_activity$p_value), ]
  
  # Add tissue column
  tf_activity$tissue <- tissue
  
  # Store in list
  all_tf_activity[[tissue]] <- tf_activity
  
  # Export individual tissue results
  output_file <- paste0("celest/output/tf_activity/TF_activity_", tissue, ".txt")
  write.table(
    tf_activity,
    output_file,
    sep = "\t",
    quote = FALSE,
    col.names = TRUE,
    row.names = FALSE
  )
  
  cat("  Saved results to:", output_file, "\n")
  cat("  Significant TFs (p < 0.05):", sum(tf_activity$p_value < 0.05), "\n")
  
  
}

cat("\n", rep("=", 60), "\n", sep = "")
cat("All tissues processed successfully!\n")
cat(rep("=", 60), "\n", sep = "")
```


## Summary Across All Tissues

```{r summary-table}
summary_table <- data.frame(
  Tissue = character(),
  Total_TFs = integer(),
  Significant_TFs = integer(),
  TFs_up = integer(),
  TFs_down = integer(),
  Percent_Significant = numeric(),
  stringsAsFactors = FALSE
)

for (tissue in names(all_tf_activity)) {
  tf_data <- all_tf_activity[[tissue]]
  n_total <- nrow(tf_data)
  sig_idx <- tf_data$p_value < 0.05
  n_sig <- sum(sig_idx)
  n_up <- sum(sig_idx & tf_data$score > 0, na.rm = TRUE)
  n_down <- sum(sig_idx & tf_data$score < 0, na.rm = TRUE)
  pct_sig <- round(100 * n_sig / n_total, 2)
  
  summary_table <- rbind(
    summary_table,
    data.frame(
      Tissue = tissue,
      Total_TFs = n_total,
      Significant_TFs = n_sig,
      TFs_up = n_up,
      TFs_down = n_down,
      Percent_Significant = pct_sig
    )
  )
}
```



# Export Results

```{r export-combined-results}
# First export the all_tf_activity list, so we don't have to do analysis again and just import
saveRDS(all_tf_activity, "celest/output/all_tf_activity.rds")
saveRDS(all_backgrounds, "celest/output/all_backgrounds.rds")

# import without running the analysis again
#all_tf_activity <- readRDS("celest/output/all_tf_activity.rds")

# Combine all tissue results into one data frame
combined_results <- do.call(rbind, all_tf_activity)

# Save combined results
write.table(
  combined_results,
  "celest/output/TF_activity_all_tissues_combined.txt",
  sep = "\t",
  quote = FALSE,
  col.names = TRUE,
  row.names = FALSE
)

# Combined results pivoted wider for tissues (each tissue has a column and the values are score) in the same sheet

combined_results_wide <- combined_results %>%

  # Keep only significant results
  filter(p_value < 0.05) %>%
  
  # Remove unwanted columns
  select(-run_id, -statistic, -condition) %>%
  
  # Pivot wider: tissue becomes columns, score becomes values
  pivot_wider(
    id_cols = c(source, wb_id, public_name, Wormbase_mode_of_regulation, Uniprot_mode_of_regulation),
    names_from = tissue,
    values_from = score
  )

write_xlsx(combined_results_wide, "celest/output/celest_combined_results_wide.xlsx")
write_xlsx(summary_table, "celest/output/celest_summary_table.xlsx")

  
# Save Excel file with separate sheets for each tissue
wb <- createWorkbook()

for (tissue in names(all_tf_activity)) {
  addWorksheet(wb, tissue)
  writeData(wb, tissue, all_tf_activity[[tissue]])
}

saveWorkbook(wb, "celest/output/TF_activity_all_tissues.xlsx", overwrite = TRUE)

cat("Results exported successfully!\n")
cat("- Combined text file: celest/output/TF_activity_all_tissues_combined.txt\n")
cat("- Excel file with all tissues: celest/output/TF_activity_all_tissues.xlsx\n")
cat("- Individual tissue files: celest/output/TF_activity_[tissue].txt\n")
```

# Overall Summary Statistics

```{r overall-summary-stats}
# Collect significant TFs with direction of activity
sig_tf_direction <- lapply(all_tf_activity, function(x) {
  x_sig <- x[x$p_value < 0.05, c("public_name", "score")]
  x_sig
})

# Combine all tissues into one data frame
sig_tf_direction_df <- do.call(rbind, sig_tf_direction)

# Add direction column
sig_tf_direction_df$direction <- ifelse(
  sig_tf_direction_df$score > 0, "Up",
  ifelse(sig_tf_direction_df$score < 0, "Down", NA)
)

# Remove any NA directions (just in case score == 0)
sig_tf_direction_df <- sig_tf_direction_df[!is.na(sig_tf_direction_df$direction), ]

# Count number of tissues per TF per direction
tf_direction_counts <- table(
  sig_tf_direction_df$public_name,
  sig_tf_direction_df$direction
)

# Convert to data frame
tf_direction_counts_df <- as.data.frame.matrix(tf_direction_counts)

# Ensure both columns exist
if (!"Up" %in% colnames(tf_direction_counts_df)) {
  tf_direction_counts_df$Up <- 0
}
if (!"Down" %in% colnames(tf_direction_counts_df)) {
  tf_direction_counts_df$Down <- 0
}

# Add total significant tissues (optional, useful for sorting)
tf_direction_counts_df$Total <- tf_direction_counts_df$Up + tf_direction_counts_df$Down

# Sort by total number of tissues
tf_direction_counts_df <- tf_direction_counts_df[
  order(tf_direction_counts_df$Total, decreasing = TRUE),
]

# Export tf_direction_counts_df to a csv file
write.csv(
  tf_direction_counts_df,
  "celest/output/TF_direction_counts_across_tissues.csv",
  row.names = TRUE
)

# Display top 20 TFs
top_tf_direction <- head(tf_direction_counts_df, 20)

cat("Top 20 TFs with Increased / Decreased Activity Across Tissues:\n")
cat("---------------------------------------------------------------\n")
for (i in seq_len(nrow(top_tf_direction))) {
  cat(sprintf(
    "%-15s: %2d tissues up, %2d tissues down (Total: %2d)\n",
    rownames(top_tf_direction)[i],
    top_tf_direction$Up[i],
    top_tf_direction$Down[i],
    top_tf_direction$Total[i]
  ))
}
```

# Session Information

```{r session-info}
sessionInfo()
```

---

## Outputs



- Individual tissue results: `celest/output/TF_activity_[tissue].txt`
- Combined results: `celest/output/TF_activity_all_tissues_combined.txt`
- Excel workbook with all tissues: `celest/output/TF_activity_all_tissues.xlsx`
- Volcano plots for each tissue: `celest/output/TF_activity_volcano_[tissue].png/.pdf`

