---
title: "TF_activity and log2FC correlation across tissues"
subtitle: ""
author: "Sharan Surya"
date: "`r Sys.Date()`"
format:
  html:
    toc: true
    toc-depth: 3
    code-fold: show
    code-tools: true
    theme: cosmo
    fig-width: 12
    fig-height: 9
    embed-resources: true
---

# Overview

To generate a plot of Uniquely differentially expressed in each tissue. Similar to Figure 4F in the manuscript, but for all targets of TFs (not just DAF-16).

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  fig.align = "center",
  dpi = 300
)

options(scipen = 10000)
```

# Load Required Packages

```{r}
#| label: setup-libraries
#| message: false
#| warning: false

library(ggplot2)
library(dplyr)
library(readxl)
library(writexl)
library(tidyr)
library(ggpmisc)

```

# Load Data

```{r load-log2fc-data}
tf_activity_log2fc <- readRDS("celest/output/tf_activity_log2fc.rds")

all_backgrounds <- readRDS("celest/output/all_backgrounds.rds")
all_tf_activity <- readRDS("celest/output/all_tf_activity.rds")

# Import and clean 50TF table - from Ryan - manually annotated..
tf_50 <- read_xlsx("celest/input/50TF_table_v2.xlsx") %>%
  select(-`TF Family`)

# Import TF annotation
tf_annotation <- read_xlsx("celest/input/Table S1 - TF annotation.xlsx", 
                           sheet = "Table S1 - TF annotation")
```


# Plot TF Activity vs Log2FC across tissues 

```{r plot-tf-activity-log2fc, fig.width=10, fig.height=6}

tf_activity_vs_log2fc <- ggplot(tf_activity_log2fc, aes(x = log2FoldChange, y = score_sig)) +
  geom_vline(xintercept = 0,
             linetype = "dashed", linewidth = 0.4, color = "grey60") +
  geom_hline(yintercept = 0,
             linetype = "dashed", linewidth = 0.4, color = "grey60") +
  # Regression on Q1 and Q3 only
  geom_smooth(
    data = ~ filter(.x, (log2FoldChange > 0 & score_sig > 0) | 
                        (log2FoldChange < 0 & score_sig < 0)),
    method = "lm", se = FALSE,
    color = "steelblue", fill = "steelblue",
    linewidth = 0.7, alpha = 0.15
  ) +
  ggpmisc::stat_poly_eq(
    data = ~ filter(.x, (log2FoldChange > 0 & score_sig > 0) | 
                        (log2FoldChange < 0 & score_sig < 0)),
    aes(label = paste(after_stat(rr.label), after_stat(p.value.label), sep = "~~")),
    formula  = y ~ x,
    parse    = TRUE,
    size     = 2.5,
    label.x  = "right",
    label.y  = "top",
    color    = "steelblue"
  ) +
  # All points
  geom_point(size = 1.8, alpha = 0.85) +
  ggrepel::geom_text_repel(
    aes(label = public_name),
    size          = 2.2,
    max.overlaps  = 15,
    segment.size  = 0.3,
    segment.color = "grey60",
    show.legend   = FALSE
  ) +
  facet_wrap(~ tissue, ncol = 5) +
  labs(
    x     = expression(bold(log[2]~FoldChange)),
    y     = "CelEsT activity score",
    title = "TF activity vs Log2FC across tissues",
  ) +
  theme_bw(base_size = 11) +
  theme(
    plot.title       = element_text(hjust = 0.5, face = "bold"),
    axis.title       = element_text(face = "bold"),
    strip.background = element_rect(fill = "grey95", color = "grey80"),
    strip.text       = element_text(size = 8, face = "bold"),
    legend.position  = "bottom",
    legend.text      = element_text(size = 8),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    plot.caption     = element_text(size = 7, color = "grey50")
  )


ggsave(
  filename = "celest/plots/tf_activity_vs_log2fc.pdf", 
  plot = tf_activity_vs_log2fc,
  width = 20, 
  height = 14, 
  units = "in",
  device = cairo_pdf
)

```

# Pull out TFs in each quadrant
Q1 (log2FoldChange > 0, score_sig > 0)
Q2 (log2FoldChange < 0, score_sig > 0)
Q3 (log2FoldChange < 0, score_sig < 0)
Q4 (log2FoldChange > 0, score_sig < 0)

```{r pull-tf-quadrants, fig.width=10, fig.height=6}
tf_activity_log2fc_quadrants <- tf_activity_log2fc %>%
  mutate(quadrant = case_when(
    log2FoldChange > 0 & score_sig > 0 ~ "Q1",
    log2FoldChange < 0 & score_sig > 0 ~ "Q2",
    log2FoldChange < 0 & score_sig < 0 ~ "Q3",
    log2FoldChange > 0 & score_sig < 0 ~ "Q4",
    TRUE ~ NA_character_
  ))

# make it wider such that columns are tissues and values are quadrants
tf_activity_log2fc_quadrants_wide <- tf_activity_log2fc_quadrants %>%
  select(public_name, wb_id, tissue, quadrant) %>%
  pivot_wider(names_from = tissue, values_from = quadrant)

# add a column which represents the quadrant across tissues for each TF. if multiple quadrants found for each TF, then concatenate them with a comma. ignore NAs

tf_activity_log2fc_quadrants_wide <- tf_activity_log2fc_quadrants_wide %>%
  rowwise() %>%
  mutate(quadrant_across_tissues = paste(na.omit(c_across(-c(public_name, wb_id))), collapse = ",")) %>%
  ungroup()
```

# Map it to phenotype table and annotation file to phenotype file (from Ryan - manually annotated).
```{r map-tf-quadrants, fig.width=10, fig.height=6}

# Add TF_family to tf_activity_log2fc_quadrants_wide by matching wb_id ~ wormbase_gene
tf_activity_log2fc_quadrants_wide <- tf_activity_log2fc_quadrants_wide %>%
  left_join(
    tf_annotation %>% select(wormbase_gene, TF_family = Family_for_FigS4),
    by = c("wb_id" = "wormbase_gene")
  )

# Add quadrant_across_tissues and TF_family to tf_50 by matching public_name
tf_50 <- tf_50 %>%
  left_join(
    tf_activity_log2fc_quadrants_wide %>% 
      select(public_name, quadrant_across_tissues, TF_family) %>% # replace with select(-wb_id) for all
      distinct(),
    by = "public_name"
  )


write_xlsx(tf_50, "celest/output/50TF_table_v2_SS.xlsx")
```


# Session Info

```{r}
sessionInfo()
```

