---
title: "Tissue - Detection and Differential Expression Analysis"
author: "Sharan Surya"
date: "`r Sys.Date()`"
format:
  html:
    toc: true
    toc-depth: 3
    code-fold: show
    code-tools: true
    theme: cosmo
    fig-width: 12
    fig-height: 9
    embed-resources: true
---

# Overview

Are tissue-specific DEGs truly tissue-specific in their expression, or are they expressed in other tissues where they're just not differentially expressed?

- **Detection:** Gene detected if normalized counts (starved OR fed) > threshold (threshold = 1)
- **DEGs:** Gene is DE if padj < 0.05 AND |log2FC| > 1



```{r setup, include=FALSE}
rm(list = ls())

knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  fig.align = "center",
  dpi = 300
)

options(scipen = 10000)

```

# Load Required Packages

```{r}
#| label: setup-libraries
#| message: false
#| warning: false

library(dplyr)
library(tidyverse)
library(ggplot2)
library(ggrepel)
library(scales) # for comma formatting in axis labels
library(viridis)
library(seriation) # for OLO reordering of hclust
library(ggpubr)

```

# Load Data

## Load TF Activity Results

Just to load tissue names.

```{r load-tf-activity}
# Load TF activity results to get tissue names
tf_activity_by_tissue <- readRDS("celest/output/all_tf_activity.rds")

tissue_names <- names(tf_activity_by_tissue)
```

## Load Log2FC Data

```{r load-log2fc-data}
# Load the log2FC file
log2fc_data <- readRDS("input/all_genes_per_tissue_log2FC_20231116.rds")

# Check unique tissues and match data with tf_activity
identical(tissue_names, unique(log2fc_data$tissue))

# Filter to protein-coding genes only
log2fc_data <- log2fc_data[log2fc_data$gene_biotype == "protein_coding", ]

```

## Load pseudobulk Data

```{r load-pseudobulk-data}
# Load the pseudobulk file
pseudobulk_data <- read_csv("input/DESeq2_pseudobulk_starv_vs_fed_20230916_minimum.csv")
```

# Data Processing

## Create Detection Matrix

```{r create-detection-matrix}
# Detection threshold
detection_threshold <- 0.1  # 10% of cells expressing a gene

# Get unique gene IDs across all tissues
all_protein_coding_gene_ids <- unique(log2fc_data$gene_id)

# Initialize detection matrix: rows = genes, cols = tissues, values = 0/1
detection_matrix <- matrix(0, 
                           nrow = length(all_protein_coding_gene_ids),
                           ncol = length(tissue_names),
                           dimnames = list(all_protein_coding_gene_ids, tissue_names))


# Fill detection matrix. If detected, TRUE = 1, FALSE = 0
for (i in 1:nrow(log2fc_data)) {
  gene <- log2fc_data$gene_id[i]
  tissue <- log2fc_data$tissue[i]
  
  # Gene detected if expressed in either condition
  pct_starved <- log2fc_data$pct.starv[i]
  pct_fed     <- log2fc_data$pct.fed[i]
  max_pct_expressed <- max(pct_starved, pct_fed, na.rm = TRUE)
  
  if (!is.na(max_pct_expressed) && max_pct_expressed >= detection_threshold) {
    detection_matrix[gene, tissue] <- 1
  }
}

genes_detected_per_tissue <- colSums(detection_matrix)

```


## Create DEG Matrix

```{r create-deg-matrix}
# Significance thresholds
padj_threshold <- 0.05
log2fc_threshold <- 0

# Filter log2fc_data to include rows where pct.starv or pct.fed is above detection threshold
log2fc_data <- log2fc_data %>%
  filter(pct.starv >= detection_threshold | pct.fed >= detection_threshold)

# Initialize DEG matrix: rows = genes, cols = tissues, values = 0/1
deg_matrix <- matrix(0,
                     nrow = length(all_protein_coding_gene_ids),
                     ncol = length(tissue_names),
                     dimnames = list(all_protein_coding_gene_ids, tissue_names))

# Fill DEG matrix
for (i in 1:nrow(log2fc_data)) {
  gene <- log2fc_data$gene_id[i]
  tissue <- log2fc_data$tissue[i]
  padj <- log2fc_data$padj[i]
  log2fc <- log2fc_data$log2FoldChange[i]
  
  if (!is.na(padj) && !is.na(log2fc)) {
    if (padj < padj_threshold && abs(log2fc) > log2fc_threshold) {
      deg_matrix[gene, tissue] <- 1
    }
  }
}

degs_detected_per_tissue <- colSums(deg_matrix)
```


# Jaccard Similarity Analysis

Calculate Jaccard statistic for each gene:

**Jaccard = (# tissues where gene is DE) / (# tissues where gene is detected)**

This answers: "Of all tissues where a gene is expressed, in what proportion is it differentially expressed?"

- Jaccard = 1.0: Gene is DE in every tissue where it's expressed
- Jaccard = 0.0: Gene is never DE despite being expressed
- Jaccard = 0.5: Gene is DE in half the tissues where it's expressed

```{r jaccard-calculation}
# Calculate Jaccard for each gene
# Jaccard = n_deg_tissues / n_detected_tissues (per gene)
calculate_jaccard <- function(detection_vector, deg_vector) {
  n_detected_tissues <- sum(detection_vector == 1)
  n_deg_tissues      <- sum(deg_vector == 1)
  
  # If gene not detected anywhere, return NA
  if (n_detected_tissues == 0) return(NA)
  
  # Jaccard = proportion of detected tissues where gene is DE
  return(n_deg_tissues / n_detected_tissues)
}

gene_jaccard_scores <- sapply(all_protein_coding_gene_ids, function(gene) {
  calculate_jaccard(detection_matrix[gene, ], deg_matrix[gene, ])
})

gene_jaccard_df <- data.frame(
  gene_id    = all_protein_coding_gene_ids,
  jaccard    = gene_jaccard_scores,
  n_detected_tissues = rowSums(detection_matrix),
  n_deg_tissues      = rowSums(deg_matrix),
  stringsAsFactors = FALSE
)

# Add gene names
gene_jaccard_df$gene_name <- log2fc_data$gene_name[match(gene_jaccard_df$gene_id, 
                                                          log2fc_data$gene_id)]

# Filter out genes not detected anywhere
gene_jaccard_detected <- gene_jaccard_df[!is.na(gene_jaccard_df$jaccard), ]

```

# Plot 1: Detection vs DEG Scatter

```{r scatter-plot, fig.width=10, fig.height=8}
tissue_detection_deg_counts <- data.frame(
  tissue   = tissue_names,
  n_detected_genes = genes_detected_per_tissue,
  n_degs   = degs_detected_per_tissue
)

detection_deg_cor <- cor.test(tissue_detection_deg_counts$n_detected_genes, 
                               tissue_detection_deg_counts$n_degs)


scatter_plot <- ggplot(tissue_detection_deg_counts, aes(x = n_detected_genes, y = n_degs)) +
  geom_point(size = 3.5, alpha = 0.75, color = "#4E6D8C") +
  geom_text_repel(
    aes(label = tissue),
    size = 2,
    box.padding = 0.5,
    max.overlaps = 50,
    segment.size = 0.25
  ) +
  geom_abline(slope = 1, intercept = 0, color = "#C0392B", linetype = "dashed", linewidth = 0.5) +
  labs(
    x = "Number of Detected Genes",
    y = "Number of Differentially Expressed Genes"
  ) +
  scale_x_continuous(labels = comma, limits = c(0, 5000)) +
  scale_y_continuous(labels = comma, limits = c(0, 5000)) +
  theme_classic(base_size = 8) +
  theme(
    text = element_text(family = "Arial"),
    axis.title = element_text(size = 8, face = "bold"),
    panel.grid = element_blank(),
    axis.title.x = element_text(margin = margin(t = 5)),
    axis.title.y = element_text(margin = margin(r = 5))

  )


ggsave(
  filename = "plots/scatter_detected_vs_deg_tissue.pdf", 
  plot = scatter_plot,
  width = 3, 
  height = 3, 
  units = "in",
  device = cairo_pdf  # <--- THIS FIXES THE FONT ERROR
)
```


# Plot 2: Jaccard Distribution
```{r jaccard-dist, fig.width=12, fig.height=8}
jaccard_median <- median(gene_jaccard_detected$jaccard)
jaccard_mean   <- mean(gene_jaccard_detected$jaccard)

jaccard_hist_plot <- ggplot(gene_jaccard_detected, aes(x = jaccard)) +
  geom_histogram(binwidth = 0.02, boundary = 0, fill = "#4E6D8C", color = "white", alpha = 0.8) +
  geom_vline(xintercept = jaccard_median, linetype = "dashed", 
             color = "#2C3E50", linewidth = 1) +
  geom_vline(xintercept = jaccard_mean, linetype = "dotted",
             color = "#16A085", linewidth = 1) +
  annotate("text", x = jaccard_median, y = Inf,
           label = paste("Median =", round(jaccard_median, 3)),
           vjust = 1.5, hjust = -0.1, size = 4, fontface = "bold",
           color = "#2C3E50") +
  annotate("text", x = jaccard_mean, y = Inf,
           label = paste("Mean =", round(jaccard_mean, 3)),
           vjust = 3, hjust = -0.1, size = 4, fontface = "bold",
           color = "#16A085") +
  scale_x_continuous(breaks = seq(0, 1, 0.1), limits = c(0, 1)) +
  scale_y_continuous(labels = comma, expand = expansion(mult = c(0, 0.05))) +
  labs(
    x = "Jaccard Index (# tissues DE / # tissues detected)",
    y = "Number of Genes",
    title = "Proportion of Detected Tissues Where Genes are Differentially Expressed",
    subtitle = "Jaccard = 1: DE in all tissues where expressed | Jaccard = 0: Never DE despite expression"
  ) +
  theme_bw(base_size = 12) +
  theme(
    plot.title    = element_text(size = 14, face = "bold", hjust = 0.5, margin = margin(b = 5)),
    plot.subtitle = element_text(size = 10, hjust = 0.5, color = "grey40", margin = margin(b = 15)),
    axis.title    = element_text(size = 11, face = "bold"),
    panel.grid = element_blank()
  )

ggsave(
  filename = "plots/jaccard_distribution_tissue.pdf", 
  plot = jaccard_hist_plot,
  width = 12, 
  height = 8, 
  units = "in",
  device = cairo_pdf  # <--- THIS FIXES THE FONT ERROR
)

```

# Plot 3: Detection vs DEG Breadth (No of tissues in both x and y)

```{r expression-vs-de-circles, fig.width=11, fig.height=10}
detection_vs_deg_breadth_plot <- ggplot(gene_jaccard_detected, aes(x = n_detected_tissues, y = n_deg_tissues)) +
  geom_count(color = "#4E6D8C", alpha = 0.65) +
  geom_abline(intercept = 0, slope = 1, linetype = "dashed", 
              color = "#C0392B", linewidth = 0.5) +
  scale_size_continuous(
    name = "Number\nof Genes",
    range = c(0.75, 5),
    labels = comma
  ) +
  labs(
    x = "Number of Tissues Gene is Detected",
    y = "Number of Tissues Gene is DE"
  ) +
  scale_x_continuous(
    limits = c(0, 18),
    breaks = seq(0, 18, 2)
  ) +
  scale_y_continuous(
    limits = c(0, 18),
    breaks = seq(0, 18, 2)
  ) +
  theme_classic(base_size = 8) +
  theme(
    text = element_text(family = "Arial"),
    axis.title    = element_text(size = 8, face = "bold"),
    legend.position  = "NA",
    axis.title.x = element_text(margin = margin(t = 5)),
    axis.title.y = element_text(margin = margin(r = 5)),
  )


ggsave(
  filename = "plots/expression_breadth_vs_de_breadth_tissue.pdf", 
  plot = detection_vs_deg_breadth_plot,
  width = 2.5, 
  height = 2.5, 
  units = "in",
  device = cairo_pdf
)
```



# Tissue Specificity

Number of gene (y) vs Number of tissues where gene is DE (x) - this shows how many genes are DE in only 1 tissue, 2 tissues, etc.

```{r tissue-specificity, fig.width=12, fig.height=8}
deg_tissue_breadth <- data.frame(
  gene_id      = rownames(deg_matrix),
  n_tissues_de = rowSums(deg_matrix)
) %>% filter(n_tissues_de > 0)

deg_breadth_barplot <- ggplot(deg_tissue_breadth, aes(x = n_tissues_de)) +
  geom_bar(fill = "#4E6D8C", color = "white", alpha = 0.8, width = 0.7) +
  #geom_text(stat = "count", aes(label = after_stat(count)), vjust = -0.5, size = 3.5) +
  scale_x_reverse(
    breaks = seq(0, 18, 2)
  ) +
  scale_y_continuous(
    trans = scales::log_trans(base = 10),
    labels = comma,
    limits = c(1, NA)
  ) +
  labs(
    x = "Number of Tissues Gene is DE",
    y = "Number of Genes",
  ) +
  theme_classic(base_size = 8) +
  theme(
    text = element_text(family = "Arial"),
    axis.title = element_text(size = 8, face = "bold"),
    axis.text.y = element_text(angle = 90, vjust = 0.5,hjust = 0.5),
    panel.grid = element_blank()
  )

ggsave(
  filename = "plots/deg_specificity_tissue.pdf", 
  plot = deg_breadth_barplot,
  width = 2.5, 
  height = 1.2, 
  units = "in",
  device = cairo_pdf
)

```


# Tissue Detection Specificity

Number of gene (y) vs Number of Tissues Gene is detected in (x) - this shows how many genes are detected in only 1 tissue, 2 tissues, etc.

```{r tissue-detection-specificity, fig.width=12, fig.height=8}
detection_tissue_breadth <- data.frame(
  gene_id      = rownames(detection_matrix),
  n_tissues_detected = rowSums(detection_matrix)
) %>% filter(n_tissues_detected > 0)

n_distinct(detection_tissue_breadth$gene_id)

detection_breadth_barplot <- ggplot(detection_tissue_breadth, aes(x = n_tissues_detected)) +
  geom_bar(fill = "#4E6D8C", color = "white", alpha = 0.8, width = 0.7) +
  #geom_text(stat = "count", aes(label = after_stat(count)), vjust = -0.5, size = 3.5) +
  scale_x_continuous(
    breaks = seq(0, 18, 2)
  ) +
  scale_y_continuous(
    trans = scales::log_trans(base = 10),
    labels = comma,
    limits = c(1, NA)
  ) +
  labs(
    x = "Number of Tissues Gene is Detected",
    y = "Number of Genes"
  ) +
  
  theme_classic(base_size = 8) +
  theme(
    text = element_text(family = "Arial"),
    axis.title = element_text(size = 8, face = "bold"),
    panel.grid = element_blank()
  )

ggsave(
  filename = "plots/detection_specificity_tissue.pdf", 
  plot = detection_breadth_barplot,
  width = 2.5, 
  height = 1.2, 
  units = "in",
  device = cairo_pdf
)
```



# Total DEGs per Tissue

Number of DEGs per tissue (y) vs Tissue (x) - this shows how many DEGs are in each tissue, regardless of whether they're unique or shared with other tissues.

```{r total-degs}
degs_per_tissue_vec <- colSums(deg_matrix)

total_deg_per_tissue <- data.frame(
  tissue = names(degs_per_tissue_vec),
  n_degs = degs_per_tissue_vec
)

total_deg_barplot <- ggplot(total_deg_per_tissue, aes(x = reorder(tissue, n_degs), y = n_degs)) +
  geom_bar(stat = "identity", fill = "#4E6D8C", alpha = 0.8) +
  geom_text(aes(label = n_degs), vjust = -0.5, size = 3.5) +
  labs(
    title = "Differentially Expressed Genes per Tissue",
    x = "Tissue",
    y = "Number of DEGs"
  ) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.1))) +
  theme_bw(base_size = 12) +
  theme(
    plot.title  = element_text(size = 14, face = "bold", hjust = 0.5),
    axis.title  = element_text(size = 12, face = "bold"),
    axis.text.x = element_text(angle = 45, hjust = 1),
    panel.grid = element_blank()
  )


ggsave(
  filename = "plots/degs_per_tissue.pdf", 
  plot = total_deg_barplot,
  width = 12, 
  height = 8, 
  units = "in",
  device = cairo_pdf
)

cat("\nTotal DEGs across all tissues:", sum(total_deg_per_tissue$n_degs), "\n")
```

# Unique DEGs per Tissue

Number of DEGs that are unique to each tissue (y) vs Tissue (x) - this shows how many DEGs are found in only one tissue, which is a stronger indication of tissue-specificity.

```{r unique-degs}
# Genes DE in exactly one tissue
single_tissue_deg_matrix <- deg_matrix[rowSums(deg_matrix) == 1, ]
single_tissue_deg_counts_per_tissue <- colSums(single_tissue_deg_matrix)

# Convert to data frame for plotting
unique_deg_per_tissue <- data.frame(
  tissue         = names(single_tissue_deg_counts_per_tissue),
  n_unique_degs  = single_tissue_deg_counts_per_tissue
)
```

```{r unique-degs-barplot, fig.width=12, fig.height=8}
unique_deg_barplot <- ggplot(unique_deg_per_tissue, aes(x = reorder(tissue, n_unique_degs), y = n_unique_degs)) +
  geom_bar(stat = "identity", fill = "#4E6D8C", alpha = 0.8) +
  geom_text(aes(label = n_unique_degs), vjust = -0.5, size = 1.5) +
  labs(
    x = "Tissue",
    y = "Number of Unique DEGs"
  ) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.1))) +
  theme_classic(base_size = 8) +
  theme(
    text = element_text(family = "Arial"),
    axis.title    = element_text(size = 8, face = "bold"),
    axis.text.x   = element_text(angle = 45, hjust = 1),
    legend.position = "none",
    panel.grid = element_blank()
  )


ggsave(
  filename = "plots/unique_degs_per_tissue.pdf", 
  plot = unique_deg_barplot,
  width = 5, 
  height = 3.5, 
  units = "in",
  device = cairo_pdf
)

cat("\nTotal unique DEGs across all tissues:", sum(unique_deg_per_tissue$n_unique_degs), "\n")
```

# Heatmaps of Log2FC per Tissue

```{r lfc-heatmap-prep}
#| eval: false

tissue_order <- c("Germline", "SGP", "hmc", "GLR",
  "Pblast", "NonStriatedMuscle", "BWM", "Excretory", 
  "Pharynx", "NonHypodermEpithelium", "Coelomocyte",
  "Glia", "NonSeamHypoderm", "Seam", "PharyngealNeuron", 
  "Neuron", "CiliatedNeuron", "Intestine")

# Up/down regulation colors
col_upregulated   <- "#7B2D8B"   # purple
col_downregulated <- "#A8C500"   # yellow-green

lfc_long <- log2fc_data %>%
  select(gene_id, gene_name, log2FoldChange, padj, tissue) %>%
  group_by(gene_name) %>%
  filter(any(padj < 0.05, na.rm = TRUE)) %>%
  ungroup()

# Build gene x tissue matrix of log2FC values
lfc_matrix <- lfc_long %>%
  select(gene_name, tissue, log2FoldChange) %>%
  pivot_wider(names_from = tissue, values_from = log2FoldChange) %>%
  column_to_rownames("gene_name") %>%
  as.matrix()

lfc_matrix <- lfc_matrix[, tissue_order[tissue_order %in% colnames(lfc_matrix)]]
lfc_matrix <- lfc_matrix[rowSums(!is.na(lfc_matrix)) > 0, ]
lfc_matrix[is.na(lfc_matrix)] <- 0

# Pearson distance + complete linkage hclust
pearson_dist_genes <- as.dist(1 - cor(t(lfc_matrix), method = "pearson"))
hclust_genes        <- hclust(pearson_dist_genes, method = "complete")

# OLO via reorder() — exactly as described in methods - takes a while!!
hclust_genes_olo <- reorder(hclust_genes, pearson_dist_genes, method = "OLO")

gene_order_levels <- hclust_genes_olo$labels[hclust_genes_olo$order]

# Consistent direction: downregulated at bottom
first_gene_mean <- mean(lfc_matrix[gene_order_levels[1], ], na.rm = TRUE)
if (first_gene_mean > 0) gene_order_levels <- rev(gene_order_levels)

lfc_long_ordered <- lfc_long %>%
  select(gene_name, tissue, log2FoldChange, padj) %>%
  complete(gene_name, tissue = tissue_order) %>%
  mutate(
    direction = case_when(
      log2FoldChange > 0 & padj < 0.05 ~ "Upregulated",
      log2FoldChange < 0 & padj < 0.05 ~ "Downregulated",
      TRUE                              ~ NA_character_
    ),
    gene_name = factor(gene_name, levels = gene_order_levels),
    tissue    = factor(tissue, levels = tissue_order)
  )

deg_heatmap_tissue <- ggplot(lfc_long_ordered, aes(x = tissue, y = gene_name, fill = direction)) +
  geom_tile(color = NA, alpha = 0.85) +
  scale_fill_manual(
    values       = c("Upregulated" = col_upregulated, "Downregulated" = col_downregulated),
    na.value     = "black",
    na.translate = FALSE,
    name         = "Direction (starved/fed)"
  ) +
  labs(
    title    = "Log2FC Across Tissues",
    subtitle = paste0(n_distinct(lfc_long$gene_id), " significant genes"),
    x        = "Tissue",
    y        = NULL
  ) +
  theme_classic(base_size = 11) +
  theme(
    plot.title       = element_text(size = 14, face = "bold", hjust = 0.5),
    plot.subtitle    = element_text(size = 11, hjust = 0.5, color = "grey40"),
    axis.text.x      = element_text(angle = 45, hjust = 1, size = 12),
    axis.text.y      = element_blank(),
    axis.ticks.y     = element_blank(),
    legend.position  = "top",
    legend.title     = element_text(size = 12, face = "bold"),
    legend.key.width = unit(2, "cm"),
    panel.grid       = element_blank(),
    panel.background = element_rect(fill = "black")
  )

ggsave(
  filename = "plots/deg_heatmap_tissue.pdf", 
  plot = deg_heatmap_tissue,
  width = 12, 
  height = 12, 
  units = "in",
  device = cairo_pdf
)

```


# Percent of TFs Significant (from CelEsT) vs Percent of Genes Detected that are DE

```{r summary-table}
summary_table <- data.frame(
  Tissue = character(),
  Total_TFs = integer(),
  Significant_TFs = integer(),
  TFs_up = integer(),
  TFs_down = integer(),
  Percent_Significant = numeric(),
  Percent_deg = numeric(),
  stringsAsFactors = FALSE
)


for (tissue in names(tf_activity_by_tissue)) {
  tf_data <- tf_activity_by_tissue[[tissue]]
  
  n_total <- nrow(tf_data)
  
  sig_idx <- tf_data$p_value < 0.05
  n_sig <- sum(sig_idx)
  
  n_up <- sum(sig_idx & tf_data$score > 0, na.rm = TRUE)
  n_down <- sum(sig_idx & tf_data$score < 0, na.rm = TRUE)
  
  pct_sig <- round(100 * n_sig / n_total, 2)
  pct_deg <- 100 * degs_detected_per_tissue[tissue] /genes_detected_per_tissue[tissue]


  summary_table <- rbind(
    summary_table,
    data.frame(
      Tissue = tissue,
      Total_TFs = n_total,
      Significant_TFs = n_sig,
      TFs_up = n_up,
      TFs_down = n_down,
      Percent_Significant = pct_sig,
      Percent_deg = pct_deg
    )
  )
}

scatter_tfs_vs_deg <- ggplot(
  summary_table,
  aes(x = Percent_deg, y = Percent_Significant)
) +
  geom_point(size = 2.5, alpha = 0.75, color = "#4E6D8C") +
  geom_smooth(method = "lm", se = FALSE, color = "#C0392B",
              linewidth = 0.6, fill = "#C0392B", alpha = 0.15) +
  geom_text_repel(
    aes(label = Tissue),
    size = 2.5,
    box.padding = 0.5,
    max.overlaps = 30
  ) +
  stat_cor(method = "pearson", label.x = 5, label.y = 95, size = 2.5) +
  stat_regline_equation(label.x = 5, label.y = 88, size = 2.5) +
  labs(
    x = "% genes detected that are DE",
    y = "% significant TFs",
  ) +
  scale_x_continuous(limits = c(0, 100)) +
  scale_y_continuous(limits = c(0, 100)) +
  theme_classic(base_size = 8) +
  theme(
    axis.title = element_text(size = 8, face = "bold"),
    panel.grid = element_blank()
  )


ggsave(
  filename = "plots/scatter_significant_tfs_vs_deg_tissue.pdf", 
  plot = scatter_tfs_vs_deg,
  width = 5, 
  height = 3.5, 
  units = "in",
  device = cairo_pdf
)

```

# Export Results

```{r export}
write.table(gene_jaccard_df, "output/gene_jaccard_results_tissue.txt", 
            sep = "\t", quote = FALSE, row.names = FALSE)
write.table(detection_matrix, "output/detection_matrix_tissue.txt",
            sep = "\t", quote = FALSE, col.names = NA)
write.table(deg_matrix, "output/deg_matrix_tissue.txt",
            sep = "\t", quote = FALSE, col.names = NA)

tissue_summary_stats <- data.frame(
  tissue             = tissue_names,
  n_detected_genes   = genes_detected_per_tissue,
  n_degs             = degs_detected_per_tissue,
  pct_deg            = round(100 * degs_detected_per_tissue / genes_detected_per_tissue, 2)
)
write.table(tissue_summary_stats, "output/summary_tissue.txt",
            sep = "\t", quote = FALSE, row.names = FALSE)

cat("Results exported\n")
```



# Session Info

```{r}
sessionInfo()
```

